A Mechanical Formalization of Higher-Ranked Polymorphic Type Inference

### 2019-05-29 sections/Discussion.tex
Discussion

This section discusses some insights that we gained from our work and contrasts
the scoping mechanisms we have employed with those in DK's algorithm.
We also discuss a way to improve the precision of their scope tracking.
Furthermore we discuss and sketch an extension of our algorithm with
an elaboration to a target calculus.

@@

\subsection{Lexically-scoped type variables}
We further extend the type system to support a simple extension, Lexically-scoped type variables.
From a practical point of view,
such extension enables the programmer to refer to type variables in a generic function signature
when implementing of that function.
For example,
$$(\lam{\lam[y] (x:a)}) : \all[a~b] a \to b \to a$$
has an inner annotation $(x:a)$ that refers to the outer generic type variable $a$.
This is not a surprising feature, since the declarative system already accepts these kind of programs
$$
\inferrule*[right=$\mathtt{DeclAnno}$]
{
    \Psi \vdash \all A \quad
    \inferrule*[right=$\mathtt{Decl\forall I}$]
    {\Psi,a \vdash e \Lto A}
    {\Psi\vdash e\Lto \all A}
}
    {\Psi\vdash (e: \all A)\To \all A}
$$

@@

In a practical point of view, programmers may not write the syntax $\Lambda a.~e : A$ directly.
The ScopedTypeVariables extension of Haskell is effective only
when the type signature is explicitly universally quantified
(the compiler translates into an expression similar to $\Lambda a.~e : A$);
otherwise the program means the normal syntax $e : \all A$,
and one may not later refer to the type variable $a$.

>>>>

This section discusses some insights that we gained from our work and contrasts
the scoping mechanisms we have employed with those in DK's algorithm.
We also discuss a way to improve the precision of their scope tracking.
Furthermore we discuss and sketch an extension of our algorithm with
an elaboration to a target calculus, and discuss an extension of our algorithm
with scoped type variables~\cite{scoped-type-variables}.

@@

\subsection{Lexically-scoped type variables}
We further extend the type system to support a simple extension,
lexically-scoped type variables~\cite{scoped-type-variables}.
We have a complete Abella formalization for this extension proving
all the metatheory discussed in the paper.

From a
practical point of view, such extension enables the programmer to
refer to type variables in a generic function signature when
implementing that function.  For example,
$$(\lam{\lam[y] (x:a)}) : \all[a~b] a \to b \to a$$
has an inner annotation $(x:a)$ that refers to the outer generic type variable $a$.
This is not a surprising feature, since the declarative system already accepts these kind of programs
$$
\inferrule*[right=$\mathtt{DeclAnno}$]
{
    \Psi \vdash \all A \quad
    \inferrule*[right=$\mathtt{Decl\forall I}$]
    {\Psi,a \vdash e \Lto A}
    {\Psi\vdash e\Lto \all A}
}
    {\Psi\vdash (e: \all A)\To \all A}
$$

@@

In practice, programmers would not write the syntax $\Lambda a.~e : A$ directly.
The ScopedTypeVariables extension of Haskell is effective only
when the type signature is explicitly universally quantified
(the compiler translates into an expression similar to $\Lambda a.~e : A$);
otherwise the program means the normal syntax $e : \all A$,
and one may not later refer to the type variable $a$.
